Bayesian regression facilitates quantitative modelling of cell metabolism

### 2023-10-31 paper/main.tex
include renders

\title[]{Bayesian regression facilitates quantitative modelling of cell
metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

A kinetic model of cellular metabolism aims to express what is known
about a cellular process in the form of an \emph{in silico}
representation of the underlying network of chemical reactions. Kinetic
models can be used to improve production of target molecules, determine
regulatory networks \citep{christodoulou_reserve_2018} and identify
potential drug targets
\citep{deberardinis_fundamentals_2016, Liberti2017}. However, the use of
kinetic models in practice is hindered by their dependence on noisy and
uncertain information sources. Quantitative in vivo measurements of
chemical abundances, and in vitro measurements relating to kinetic
parameters, both contain vital information but are notoriously
inaccurate (REFERENCES). Practically useful kinetic modelling therefore
requires a principled statistical approach that encompasses multiple
possible model parameterisations.

>>>>

\title[]{Bayesian regression facilitates quantitative modelling of cell
metabolism}
\makeatletter
\begin{document}
\maketitle
\begin{abstract}
This paper presents Maud, a command-line application that implements
Bayesian statistical inference for kinetic models of biochemical
metabolic reaction networks. Maud takes into account quantitative
information from omics experiments and background knowledge, as well as
structural information about kinetic mechanisms, regulatory interactions
and enzyme knockouts. Our paper reviews the existing options in this
area, presents a case study illustrating how Maud can be used to analyse
a metabolic network and explains the biological, statistical and
computational design decisions underpinning Maud.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, enhanced, sharp corners, breakable]}{\end{tcolorbox}}\fi

@@

A kinetic model of cellular metabolism aims to express what is known
about a cellular process in the form of an \emph{in silico}
representation of the underlying network of chemical reactions. Kinetic
models can be used to improve production of target molecules, determine
regulatory networks \citep{christodoulou_reserve_2018} and identify
potential drug targets
\citep{deberardinis_fundamentals_2016, Liberti2017}. However, the use of
kinetic models in practice is hindered by their dependence on noisy and
uncertain information sources. Quantitative in vivo measurements of
chemical abundances, and in vitro measurements relating to kinetic
parameters, both contain vital information but are notoriously
inaccurate {[}siskos2017; tabb2010{]}. Practically useful kinetic
modelling therefore requires a principled statistical approach that
encompasses multiple possible model parameterisations.
